---
title: "Query GGBN"
author: "Michael JV O'Mahoney"
date: "February 2021"
output: html_notebook
---

## Loading Libraries & Data

* Load in the required libraries
```{r}
# needed to run fromJSON(), 
library('jsonlite')
# needed to run rbind.fill()
library('plyr')
# needed to run the progress_bar
library('progress')
# needed to run flatten()
library('purrr')
# needed to run RCurl()
library('RCurl')
# needed to run str_replace_all()
library('stringr')
# needed to run write_tsv()
library('readr')
```

#

* Read in the Gustav Paulay edit of Kozloff's List of Marine Inverts in Puget Sound and remove the headers from the resulting data frame.
```{r}
# Load in the data containing the taxonomy to query
data <- read.csv('GP_kozloff_edits.csv', header = TRUE, sep = ",")
# Remove the header from "data"
names(data) <- NULL
```


## Preparing the Input

* Store the base URL used to query the API
```{r}
base <- "http://data.ggbn.org/ggbn_portal/api/search?getSampletype&name="
```

* Extract column containing taxonomic names from `data` and store it as `taxa`
```{r}
## Note the double brackets "[[]]" in "data[[2]]" as it extracts the column 
## from the data frame and returns it as a vector. You could also write it as 
## "data[,2]" and get the same result. Simply writing "data[2]" returns a 
## a sublist of the "data" with the class data.frame.
taxa <- data[[2]]

# trims off leading and trailing whitespace
taxa <- trimws(taxa)
```

* Create an empty data frame to add the query output to
```{r}
result <- data.frame()
```

* Create new instance of `progress_bar` with ETA
```{r}
pb <- progress_bar$new(
  format = "  downloading [:bar] :percent eta: :eta",
  total = length(taxa), clear = FALSE, width= 60)
```

## Running the Loop

* This builds the URL, makes GET request, and uses `rbind.fill()` to append the `request` variable to the `result` data frame and fill in null values with NA
  * the `purrr::flatten()` function allows null results to be coerced into data frames

```{r}
for (i in 1:length(taxa)) {
  # Initiate progress bar with ETA
    pb$tick(0)
    pb$tick()

    request <- flatten(fromJSON(getURL(URLencode(paste0(base,taxa[i])))))
    result <- rbind.fill(result, as.data.frame(request))
}
```


## Cleaning up the Result

* Remove "fullScientificName_nc=" which is appended to each species value as an artifact of the query
```{r}
result$filters <- str_replace_all(result$filters, "^.*=", "")
```

## Saving the Result

* Write `result` to TSV
```{r}
write_tsv(result, 'GGBN_Query_results.tsv', na = "NA")
```
